An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-09-23 nips_2016.tex

{\color{blue} Daniel: Hold on, the following Lemma seems to add $X_{\rm norm}$
to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
didn't the previous section already justify that $\Delta^* \approx \Delta +
X_{\rm norm}$? That is, I don't think we need to add another $X_{\rm norm}$; I
think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
match what we wrote earlier.}

@@

\begin{proof}
We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
\begin{equation}
    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
\end{equation}
and convolve with the distribution of $X_n$ {\color{blue} (Daniel: to be clear I
think this was supposed to be $X_{\rm norm}$ but see comments earlier about why
I think it is not necessary.)} which has density $\phi(X/\sigma_n)$ where
$\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of $P$ and
$Q$
\begin{equation}
    P'(y) = {\rm Pr}(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
\end{equation}
Now we convolve with the distribution of $X_{\rm corr}$ which gives:
\begin{equation}
    P''(y) = {\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
\end{equation}
Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
norm}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
log} > 0)$ by at most $\epsilon$.
\end{proof}

@@

The deconvolution approach can be used to construct a correction variable
$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
saved offline. As we have shown above, errors in the ``limit'' function
propagate directly through as errors in the acceptance test.  To achieve a test
error of say $1e-6$ (close to single floating point precision), we need a $y'$
spacing of $1e-6$. It should not be necessary to tabulate values all the way to
$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
resolution is 10,000, this leads to a table with about 1 billion values, which
can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
in $y$, it should be possible to achieve similar accuracy with a much smaller
table. We leave further analysis and experiments with $g(x,y)$ as future work.
{\color{blue} Daniel: I will look at the code to understand this paragraph better.}

@@

\begin{itemize}[noitemsep]
    \item The proof of Lemma~\ref{lem:worst_case}.
    \item More information on Section~\ref{ssec:gaussians}.
    \item More information on Section~\ref{ssec:logistic}.
    %\item More information on Section~\ref{ssec:nets}. % Daniel: this is in
    %progress.
\end{itemize}

>>>>

{\color{blue} Daniel: Hold on, the following Corollary seems to add $X_{\rm norm}$
to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
didn't the previous section already justify that $\Delta^* \approx \Delta +
X_{\rm norm}$? That is, I don't think we need to add another $X_{\rm norm}$; I
think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
match what we wrote earlier.

I think the easiest way to resolve this is to do the following: we assume that
$\Delta^*$ is already $\Delta + X_{\rm noise}$. Then all we need to do is add
$X_{\rm corr}$. So this is the same as what we are doing, except we do not need
to apply Lemma~\ref{lem:cdf_bounds} twice; we just need to do it once because we
are only adding $X_{\rm corr}$, not $X_{\rm norm}$.}

@@

\begin{proof}
We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
\begin{equation}
    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
\end{equation}
and convolve with the distribution of $X_n$ {\color{blue} (Daniel: to be clear I
think this was supposed to be $X_{\rm norm}$ instead of $X_n$ but see comments
earlier about why I think it is not necessary.)} which has density
$\phi(X/\sigma_n)$ where $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the
next iteration of $P$ and $Q$:
\begin{equation}
    P'(y) = {\rm Pr}(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
\end{equation}
Now we convolve with the distribution of $X_{\rm corr}$ which gives:
\begin{equation}
    P''(y) = {\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
\end{equation}
Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
norm}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
log} > 0)$ by at most $\epsilon$.
\end{proof}

@@

The deconvolution approach can be used to construct a correction variable
$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
saved offline. As we have shown above, errors in the ``limit'' function
propagate directly through as errors in the acceptance test.  To achieve a test
error of say $1e-6$ (close to single floating point precision), we need a $y'$
spacing of $1e-6$. It should not be necessary to tabulate values all the way to
$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
resolution is 10,000, this leads to a table with about 1 billion values, which
can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
in $y$, it should be possible to achieve similar accuracy with a much smaller
table. We leave further analysis and experiments with $g(x,y)$ as future work.
{\color{blue} Daniel: I will look at the code to understand this paragraph
better because right now I am confused.}

@@

\begin{itemize}[noitemsep]
    \item The proof of Lemma~\ref{lem:worst_case}.
    \item More information on Section~\ref{ssec:gaussians}.
    \item More information on Section~\ref{ssec:logistic}.
\end{itemize}

{\color{blue} Daniel: Note that everything after
Appendix~\ref{app:worst_case_proof} has yet to be revised. I don't know what we
will include in them.}
